Sparse Matrix Vector Multiplication with Multiple vectors and Multiple Matrices on the MIC Architecture

### 2013-10-16 ieee_paper.tex
more citations

\cite{Im01, Saad94sparskit} \cite{DBLP:journals/ijhpca/ShantharamCR11,
  DBLP:journals/corr/abs-1101-0091 }

The multiplication of sparse matrix by a dense vector (SpMV) is an
important kernel in many fields from scientific computing~\cite{} to
graph analysis (recommender systems~\cite{Brin98}, graph
drawing~\cite{Koren05}). So improving the performance of SpMV has
captured the interest of many researcher; including the development of
various implemetation for multi-core
CPUs~\cite{Buluc2009SPAA,Williams07} and
GPUs~\cite{Bell08, conf/ipps/KreutzerHWFBB12,
  journals/concurrency/VazquezFG11,kumar2012accelerating}. The main
challenge that is faced to obtain good performance for SpMV is that
the operation are conducted using memory location that are irregular
and often unpredictible. That make that the kernels are mostly memory
bound and there is a significant instruction overhead per flop.

@@

An other solution, and the one we pursue in this paper, is to compute
multiple SpMV at once on matrices that have the same sparsity
patterns. Obviously not all the applications have such a property. But
important classes of applications such as graph
recommendation~\cite{}, eigensolving~\cite{} and the computation of
derivative of Radial Basis Functions(RBF)~\cite{} can use multiple
SpMVs simultaneously. In this paper in particular, we investigate the
case of the derivative of RBFs where four derivatives of four
different function is expressed as the multiplication of four vectors
by four matrices with identical sparsity patterns leading to the
simultaneous execution of 16 SpMVs at a time.

>>>>

\cite{DBLP:journals/ijhpca/ShantharamCR11,
  DBLP:journals/corr/abs-1101-0091 }

The multiplication of sparse matrix by a dense vector (SpMV) is an
important kernel in many fields from scientific computing~\cite{} to
graph analysis (recommender systems~\cite{Brin98}, graph
drawing~\cite{Koren05}). So improving the performance of SpMV has
captured the interest of many researcher; including the development of
various implementations for multi-core
CPUs~\cite{Buluc2009SPAA,Williams07} and
GPUs~\cite{Bell08, conf/ipps/KreutzerHWFBB12,
  journals/concurrency/VazquezFG11,kumar2012accelerating}. The main
challenge that is faced to obtain good performance for SpMV is that
the operation are conducted using memory location that are irregular
and often unpredictible. That make that the kernels are mostly memory
bound and there is a significant instruction overhead per flop.

@@

An other solution, and the one we pursue in this paper, is to compute
multiple SpMV at once on matrices that have the same sparsity
patterns. Obviously not all the applications have such a property. But
important classes of applications such as graph
recommendation~\cite{}, eigensolving~\cite{} and the computation of
derivative of Radial Basis Functions(RBF)~\cite{} can use multiple
SpMVs simultaneously. In this paper in particular, we investigate the
case of the derivative of RBFs where four derivatives of four
different function is expressed as the multiplication of four vectors
by four matrices with identical sparsity patterns leading to the
simultaneous execution of 16 SpMVs at a time. Using multiple vectors
at a time has been investigated before to densify the
computations~\cite{Im01} but this is the first time that the
densification comes from adding both vectors and matrices.
